Subject-level developmental relationships

### 2024-08-28 3_developmental_relationships.qmd
3

```{r}
d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         rawscore, percentile) |>
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", names
```

```{r}
d_sub |>
  left_join(d_cdi)
```

@@

OK, let's take a look at these subject and word-level measures.

```{r}

acc_rt <- inner_join(ms_acc, ms_rt) |>
  mutate(age_group_years = cut(age, breaks = c(12,24,36,48,60)))

head(acc_rt)
```

@@

```{r}
GGally::ggpairs(select(ungroup(acc_rt), -dataset_name, -age_group_years, -n_accuracy, -n_rt, -administration_id), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

```
It turns out that the correlations between elogit(accuracy), log(age), and log(rt) are all stronger. So we'll adopt these.

```{r}
GGally::ggpairs(select(ungroup(acc_rt), elogit, log_rt, log_age), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

So these are somewhat related to one another. Let's take a deeper look.

```{r}
ggplot(acc_rt, aes(x = log_rt, y = elogit, col = log_age)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = .5, lty = 2) + 
  xlab("Mean log RT") + 
  ylab("Mean elogit(accuracy)")
```

Broken out by age group.

```{r}
ggplot(acc_rt, aes(x = log_rt, y = elogit)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = .5, lty = 2) +
  facet_wrap(~age_group_years) + 
  xlab("Mean log RT") + 
  ylab("Mean elogit accuracy")

@@

By dataset.

```{r}
ggplot(acc_rt, aes(x = log_rt, y = elogit, col = log_age)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = .5, lty = 2) +
  facet_wrap(~dataset_name) + 
  xlab("Mean log RT") + 
  ylab("Mean elogit accuracy")

@@

```{r}
acc_rt_mat <- acc_rt |>
  ungroup() |>
  select(log_age, elogit, log_rt) |>
  mutate(log_age = scale(log_age)[,1], 
         log_rt = scale(log_rt)[,1],
         elogit = scale(elogit)[,1] )|>
  as.matrix()

acc_rt_prc <- prcomp(acc_rt_mat, 2)

>>>>

This is broken due to some repeated fields.

```{r}
d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         rawscore) |>
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", values_from = "rawscore", 
              values_fill = NA, values_fn = mean)
```

```{r}
d_sub_cdi <- d_sub |>
  left_join(d_cdi) |>
  mutate(log_age = log(age)) |>
  ungroup()
```

@@

@@

```{r}
GGally::ggpairs(select(d_sub_cdi, age, log_age, rt, log_rt, 
                       short_window_accuracy, long_window_accuracy, 
                       short_window_elogit, long_window_elogit, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

Let's select the best of these.

```{r}
GGally::ggpairs(select(d_sub_cdi, log_age, rt,  
                       long_window_accuracy, prod, comp), 
        progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

<!-- So these are somewhat related to one another. Let's take a deeper look.  -->

<!-- ```{r} -->
<!-- ggplot(acc_rt, aes(x = log_rt, y = elogit, col = log_age)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_smooth(method = "lm") + -->
<!--   geom_hline(yintercept = .5, lty = 2) +  -->
<!--   xlab("Mean log RT") +  -->
<!--   ylab("Mean elogit(accuracy)") -->
<!-- ``` -->

<!-- Broken out by age group.  -->

<!-- ```{r} -->
<!-- ggplot(acc_rt, aes(x = log_rt, y = elogit)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_smooth(method = "lm") + -->
<!--   geom_hline(yintercept = .5, lty = 2) + -->
<!--   facet_wrap(~age_group_years) +  -->
<!--   xlab("Mean log RT") +  -->
<!--   ylab("Mean elogit accuracy") -->

<!-- ``` -->

<!-- By dataset.  -->

<!-- ```{r} -->
<!-- ggplot(acc_rt, aes(x = log_rt, y = elogit, col = log_age)) + -->
<!--   geom_point(alpha = .3) + -->
<!--   geom_smooth(method = "lm") + -->
<!--   geom_hline(yintercept = .5, lty = 2) + -->
<!--   facet_wrap(~dataset_name) +  -->
<!--   xlab("Mean log RT") +  -->
<!--   ylab("Mean elogit accuracy") -->

<!-- ``` -->

@@

```{r}
d_sub_cdi_mat <- d_sub_cdi |>
  ungroup() |>
  select(log_age, rt,  long_window_accuracy, prod) |>
  mutate(log_age = scale(log_age)[,1], 
         rt = scale(rt)[,1],
         long_window_accuracy = scale(long_window_accuracy)[,1], 
         prod = scale(prod)[,1])

d_sub_cdi_mat <- d_sub_cdi_mat |>
  filter(complete.cases(d_sub_cdi_mat)) |>
  as.matrix()

acc_rt_prc <- prcomp(d_sub_cdi_mat, 2)
